---
params: 
  set_title:
    label: "report title"
    value: "Project"
    input: text
  set_author: 
    label: "author"
    value: "Miles Smith"
    input: text
title: "`r params$set_title`"
author: "`r params$set_author`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    highlight: "pygments"
    toc: TRUE
    toc_depth: 3
    toc_float: TRUE
  pdf_document:
    highlight: "pygments"
    toc: TRUE
    toc_depth: 3
always_allow_html: true
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 1500px !important;
      width: 1500px !important;
    }
    body {
      max-width: 1500px !important;
    }
    caption {
      color: black;
      font-weight: bold;
      font-size: 1.0em;
    }
```

```{r setup, include=FALSE}
options(width = 1200)
knitr::opts_chunk$set(echo = FALSE, root.dir = here::here(), warning=FALSE, message=FALSE)
ggplot2::theme_set(ggpubr::theme_pubr())
comparison_grouping_variable <- params[["comparison_grouping_variable"]]
diffused_comparison <- rlang::sym(comparison_grouping_variable)
diffused_comparison <- rlang::enquo(diffused_comparison)
source("code/plan/14_report.funcs.R")
```

```{r}
targets::tar_load(dataset_with_scores)
targets::tar_load(res)
```

```{r study_group_table}
md <- getMetaData(dataset_with_scores)
if (knitr::is_html_output()){
  md %>%
    janitor::tabyl(comparison_grouping_variable) %>%
    janitor::adorn_pct_formatting(digits = 2) %>%
    knitr::kable() %>%
    kableExtra::kable_styling(
      bootstrap_options =
        c(
          "striped",
          "hover",
          "condensed",
          "responsive"
          ),
      full_width = FALSE
      ) %>%
    kableExtra::column_spec(
      1,
      bold = TRUE,
      border_right = TRUE,
      width = "10em"
      ) %>%
    kableExtra::column_spec(
      2,
      border_right = TRUE,
      width = "10em"
      ) %>%
    kableExtra::column_spec(
      3,
      border_right = FALSE,
      width = "10em"
      )
} else if(knitr::is_latex_output()){
  md %>%
    janitor::tabyl(comparison_grouping_variable) %>%
    dplyr::rename(
      `Disease classification` = comparison_grouping_variable,
      Number = n,
      Percent = percent
      ) %>%
    flextable::qflextable() %>%
    flextable::set_formatter(
      Percent = function(x) sprintf("%.01f", x*100)
    )
}

```

\newpage

# Differential Gene Expression {.tabset .tabset-fade .tabset-pills}

```{r DE_table}
results_tbl <-
  res %>% 
    # purrr::map(alt_summary, lfcThreshold = params[["lfcThreshold"]]) %>%
  purrr::map(alt_summary, lfcThreshold = 0.1) %>%
    tibble::enframe() %>%
    tidyr::unnest(cols = c(value)) %>%
    dplyr::mutate(
      `up %` = round(
        100*(as.integer(up)/as.integer(total)),
        digits = 2
        ),
      `down %` = round(
        100*(as.integer(down)/as.integer(total)),
        digits = 2
        ),
      `low counts %` = round(
        100*(as.integer(lowcounts)/as.integer(total)),
        digits = 2
        ),
      name = stringr::str_replace_all(
        string = name, 
        pattern = "_",
        replacement = " "
        )
    )

if (knitr::is_html_output()){
  dplyr::select(
    .data = results_tbl,
    comparison = name,
    up,
    `up %`,
    down,
    `down %`,
    `mean count cutoff` = ft,
    `low counts` = lowcounts,
    `low counts %`
  ) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(
    bootstrap_options =
      c(
        "striped",
        "hover",
        "condensed",
        "responsive"
        )
    ) %>%
  kableExtra::footnote(
    paste(
      alt_summary(res[[1]])$total,
      "with nonzero total read count", "\nOutliers: ",
      alt_summary(res[[1]])$outlier)
    )
} else if (knitr::is_latex_output()){
  dplyr::select(
    .data = results_tbl,
    Comparison = name,
    `# Up\nregulated` = up,
    `% Up\nregulated` = `up %`,
    `# Down\nregulated` = down,
    `% Down\nregulated` = `down %`,
    `Mean count\ncutoff` = ft,
    `# Low\ncounts` = lowcounts,
    `% Low\ncounts` = `low counts %`
    ) %>%
  flextable::qflextable() %>%
  flextable::set_formatter(
    `% Up\nregulated` = \(x) sprintf("%.02f", x),
    `% Down\nregulated` = \(x) sprintf("%.02f", x),
    `% Low\ncounts` = \(x) sprintf("%.02f", x)
  ) %>%
  flextable::bold(part = "header") %>%
  flextable::align(
    align = "left",
    part = "body",
    j = 1
    )
  }
```

\newpage

## 25 genes with largest negative log-fold change in experimental group {.tabset .tabset-fade .tabset-pills}

```{r 25_top_down_DE, echo = FALSE}
targets::tar_load(down_tables)
```

```{r 25_top_down_DE_tbl, echo = FALSE, results='asis'}
if (knitr::is_html_output()){
  purrr::iwalk(
    .x = down_tables, 
    .f = \(dtbl, dtbl_name)
    process_deg_kable(
        deg_table = dtbl,
        deg_table_name = dtbl_name,
        .color = "tomato",
        .direction = "down"
      ) %>%
      print()
  )
} else if (knitr::is_latex_output()){
  purrr::iwalk(
    .x = down_tables,
    .f = \(dtbl, dtbl_name) 
      process_deg_flextable(dtbl, dtbl_name) %>%
      knitr::knit_print() %>%
      cat()
  )
}
```

## 25 genes with largest positive log-fold change in experimental group {.tabset .tabset-fade .tabset-pills}

```{r 25_top_up_DE, echo = FALSE}
targets::tar_load(up_tables)
```

```{r 25_top_up_DE_tbl, echo = FALSE, results='asis'}
if (knitr::is_html_output()){
  purrr::iwalk(
    .x = up_tables,
    .f = \(utbl, utbl_name)
      process_deg_kable(
        deg_table = utbl,
        deg_table_name = utbl_name,
        .color = "lightgreen",
        .direction = "up"
      ) %>%
      print()
  )
} else if (knitr::is_latex_output()){
  purrr::iwalk(
    .x = up_tables,
    .f = \(utbl, utbl_name)
      process_deg_flextable(utbl, utbl_name) %>%
      knitr::knit_print() %>% 
      cat()
  )
}
```

\newpage

## Expression of top class DE genes by class {.tabset .tabset-fade .tabset-pills}

```{r}
targets::tar_load(deg_class)
targets::tar_load(final_md)
targets::tar_load(vsc_exprs)
targets::tar_load(annotation_info)
targets::tar_load(group_pal)
targets::tar_load(up_tables)
targets::tar_load(down_tables)
```

### Sorted by responder status

```{r top_class_DE_disease_sort, fig.width=12, fig.height=9}
deg_class_tbl <-
  tibble::as_tibble(
    deg_class,
    rownames = "gene"
    )

comparisons = unique(deg_class_tbl[["comparison"]])

final_md_tbl <-
  final_md %>%
  tibble::as_tibble(rownames = "subject") %>%
  dplyr::filter(subject %in% colnames(vsc_exprs))

range_bound = 4
zscore_range <- c(-range_bound:range_bound)
comparison_names <- union(names(up_tables), names(down_tables))

for (i in comparison_names){
  #comparison_genes <- dplyr::filter(deg_class_tbl, comparison == i) %>% dplyr::pull(gene)
  comparison_genes <-
    dplyr::pull(
      dplyr::bind_rows(
        up_tables[[i]],
        down_tables[[i]]),
      gene
      )
  
  people <-
    dplyr::filter(
      .data = final_md_tbl,
      {{diffused_comparison}} %in% stringr::str_split(i, "_vs_", simplify = TRUE)) %>%
    dplyr::pull(subject)
  
  matrix_to_plot <- t(vsc_exprs)[people,comparison_genes]
  
  sample_disease_order_info <-
    annotation_info %>%
    tibble::as_tibble(rownames = "subject") %>%
    dplyr::filter(subject %in% people) %>%
    dplyr::arrange({{diffused_comparison}})
  
  sample_disease_order <-
    sample_disease_order_info %>%
    dplyr::pull(subject)
  
  sample_disease_order_count <-
    sample_disease_order_info %>%
    dplyr::count({{diffused_comparison}}) %>%
    dplyr::pull(n) %>%
    purrr::accumulate(sum)
    
  matrix_to_plot[
    sample_disease_order,
    matrixStats::colVars(matrix_to_plot) > 1e-10
    ] %>% 
    scale() %>%
      pheatmap::pheatmap(
        #scale = 'column',
        main = i,
        fontsize = 9,
        fontsize_col = 9,
        cluster_rows = FALSE,
        show_rownames = FALSE,
        annotation_row = annotation_info[sample_disease_order, ],
        annotation_colors = group_pal,
        angle_col = "315",
        breaks = zscore_range,
        color = viridisLite::viridis(n = length(zscore_range), option = "D"),
        border_color = NA,
        gaps_row = sample_disease_order_count
      )
  }
```

### Sorted by cluster

```{r top_class_DE_cluster_sort, fig.width=12, fig.height=9}

deg_class_tbl <- as_tibble(deg_class, rownames = "gene")
comparisons = unique(deg_class_tbl[["comparison"]])

final_md_tbl <- final_md %>% as_tibble(rownames = "subject") %>% dplyr::filter(subject %in% colnames(vsc_exprs))

range_bound = 4
zscore_range <- c(-range_bound:range_bound)
comparison_names <- union(names(up_tables), names(down_tables))

for (i in comparison_names){
  #comparison_genes <- dplyr::filter(deg_class_tbl, comparison == i) %>% dplyr::pull(gene)
  comparison_genes <- dplyr::pull(bind_rows(up_tables[[i]], down_tables[[i]]), gene)
  
  people <-
    dplyr::filter(
      .data = final_md_tbl,
      {{diffused_comparison}} %in% stringr::str_split(i, "_vs_", simplify = TRUE)) %>%
    dplyr::pull(subject)
  
  matrix_to_plot <- t(vsc_exprs)[people,comparison_genes]
  
  sample_cluster_order_info <-
    annotation_info %>%
    as_tibble(rownames = "subject") %>%
    dplyr::filter(subject %in% people) %>%
    dplyr::arrange(cluster)
  
  sample_cluster_order <-
    sample_cluster_order_info %>%
    dplyr::pull(subject)
  
  sample_cluster_order_count <-
    sample_cluster_order_info %>%
    dplyr::count(cluster) %>%
    dplyr::pull(n) %>%
    purrr::accumulate(sum)
  
  matrix_to_plot[
    sample_cluster_order,
    colVars(matrix_to_plot) > 1e-10
    ] %>% 
      pheatmap::pheatmap(
        scale = 'column',
        main = i,
        fontsize = 9,
        fontsize_col = 9,
        cluster_rows = FALSE,
        show_rownames = FALSE,
        annotation_row = annotation_info,
        annotation_colors = group_pal,
        angle_col = 315,
        breaks = zscore_range,
        color = viridis(n = length(zscore_range), option = "D"),
        border_color = NA,
        gaps_row = sample_cluster_order_count
      )
  }
```

### Sorted by hierarchical clustering

```{r top_class_DE_hierarchical_sort, fig.width=12, fig.height=9}
deg_class_tbl <- as_tibble(deg_class, rownames = "gene")
comparisons = unique(deg_class_tbl$comparison)

final_md_tbl <- final_md |> tibble::as_tibble(rownames = "subject") |> dplyr::filter(subject %in% colnames(vsc_exprs))

range_bound = 4
zscore_range <- c(-range_bound:range_bound)
comparison_names <- union(names(up_tables), names(down_tables))

for (i in comparison_names){
  #comparison_genes <- dplyr::filter(deg_class_tbl, comparison == i) %>% dplyr::pull(gene)
  comparison_genes <- dplyr::pull(bind_rows(up_tables[[i]], down_tables[[i]]), gene)
  
  people <-
    dplyr::filter(
      .data = final_md_tbl,
      {{diffused_comparison}} %in% stringr::str_split(i, "_vs_", simplify = TRUE)) %>%
    dplyr::pull(subject)
  
  matrix_to_plot <- t(vsc_exprs)[people,comparison_genes]
  
  matrix_to_plot[,
    colVars(matrix_to_plot) > 1e-10
    ] %>% 
      pheatmap(
        scale = 'column',
        main = i,
        fontsize = 9,
        fontsize_col = 9,
        cluster_rows = TRUE,
        show_rownames = FALSE,
        annotation_row = annotation_info,
        annotation_colors = group_pal,
        angle_col = 315,
        breaks = zscore_range,
        color = viridis(n = length(zscore_range), option = "D"),
        border_color = NA
      )
  }
```

\newpage

## Expression of all top DE genes {.tabset .tabset-fade .tabset-pills}

```{r}
tar_load(deg_means)
```

```{r Expression_of_all_top_DE_genes, fig.width=12, fig.height=9}

#### sort by disease ####
sample_disease_order_info <- 
  annotation_info %>%
  as_tibble(rownames = "subject") %>%
  dplyr::arrange({{diffused_comparison}})

sample_disease_order <-
  sample_disease_order_info %>%
  dplyr::pull(subject)

sample_disease_order_count <-
  sample_disease_order_info %>%
  dplyr::count({{diffused_comparison}}) %>%
  dplyr::pull(n) %>%
  purrr::accumulate(sum)


range_bound = 4
zscore_range <- c(-range_bound:range_bound)

if (length(colnames(deg_means)) >= 2){
column_cluster <- TRUE
} else {
column_cluster <- FALSE
}
```

### Sorted by responder status

```{r Expression_of_all_top_DE_genes_disease_sort, fig.width=12, fig.height=9}
t(vsc_exprs)[sample_disease_order,colnames(deg_means)] %>% 
    pheatmap(
      scale = 'column',
      fontsize = 9,
      fontsize_col = 6,
      show_rownames = FALSE,
      annotation_row = annotation_info,
      annotation_colors = group_pal,
      angle_col = 315,
      breaks = zscore_range,
      cluster_rows = FALSE,
      cluster_cols = column_cluster,
      color = viridis(n = length(zscore_range)-1, option = "D"),
      gaps_row = sample_disease_order_count,
      border_color = NA
      )
```

### Sorted by hierarchical clustering

```{r Expression_of_all_top_DE_genes_auto_sort, fig.width=12, fig.height=9}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)


t(vsc_exprs)[,colnames(deg_means)] %>% 
    pheatmap(
      scale = 'column',
      fontsize = 9,
      fontsize_col = 6,
      show_rownames = FALSE,
      annotation_row = annotation_info,
      annotation_colors = group_pal,
      angle_col = 315,
      breaks = zscore_range,
      cluster_rows = TRUE,
      cluster_cols = column_cluster,
      color = viridis(n = length(zscore_range)-1, option = "D"),
      border_color = NA
      )
```

\newpage

## Volcano plot of gene expression

```{r Volcano_plot_of_gene_expression, fig.height=6, fig.width=9}

inverse_format <- function() {
  function(x) format(-x,) 
}

for(k in seq_along(res)){

    volcano_data <- 
      res[[k]] %>%
      as_tibble(rownames = "gene") %>%
      dplyr::mutate(sig = case_when(
        padj > 0.01 ~ "insig",
        padj < 0.01 & log2FoldChange <= -0.25 ~ "down",
        padj < 0.01 & log2FoldChange >= 0.25 ~ "up",
        padj < 0.01 & between(log2FoldChange, -1, 1) ~ "ignore",
        is.na(padj) ~ "insig")
      )

    volcano_data_up =
      top_n(
        x = dplyr::filter(
          .data = volcano_data,
          sig == "up"
          ),
        n = 20, 
        wt = log2FoldChange
        ) %>%
      dplyr::mutate(
        sig = sig %>%
          as_factor() %>%
          fct_recode(
            "Up-regulated" = "up"
            )
        )
    volcano_data_down = 
        top_n(
        x = dplyr::filter(
          .data = volcano_data,
          sig == "down"
          ),
        n = -20,
        wt = log2FoldChange
        ) %>%
      dplyr::mutate(
        sig = sig %>%
          as_factor() %>%
          fct_recode(
            "Down-regulated" = "down"
            )
        )

  vp <-
    volcano_data %>%
      dplyr::mutate(
        sig = sig %>%
          as_factor() %>%
          fct_recode(
            "p > 0.05" = "insig",
            "Up-regulated" = "up",
            "Down-regulated" = "down"
            )
        ) %>%
      dplyr::filter(sig != "ignore") %>%
      ggplot(
        aes(
          x = log2FoldChange,
          y = (1/padj),
          color = sig
        ),
        alpha = 0.25
      ) +
      geom_point() +
      geom_text_repel(
        data = volcano_data_up,
        aes(
          x = log2FoldChange,
          y = 1/padj,
          label = gene
          ),
        show.legend = FALSE
      ) +
      geom_text_repel(
        data = volcano_data_down,
        aes(
          x = log2FoldChange,
          y = 1/padj,
          label = gene
        ),
        show.legend = FALSE
      ) +
      scale_y_log10(
        labels = inverse_format()
      ) +
      scale_color_manual(values = c('grey',"blue",'red')) +
      labs(x = "Log 2 fold change",
           y = "Adjusted p-value",
           title = names(res)[[k]])

  print(vp)
}
```

\newpage

## Expression of selected interferon-stimulated genes {.tabset .tabset-fade .tabset-pills}

```{r Expression_of_selected_interferon-stimulated genes, fig.height=6, fig.width=9}
tar_load(ISGs)
```

### Sorted by responder status

```{r Expression_of_ISGs_disease_sort, fig.width=12, fig.height=9}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
t(vsc_exprs)[sample_disease_order,ISGs] %>%
    pheatmap(
      scale = 'column',
      fontsize = 8,
      show_rownames = FALSE,
      annotation_row = annotation_info,
      annotation_colors = group_pal,
      angle_col = 315,
      breaks = zscore_range,
      cluster_rows = FALSE,
      color = viridis(n = length(zscore_range)-1, option = "D"),
      gaps_row = sample_disease_order_count,
      border_color = NA)
```

### Sorted by hierarchical clustering

```{r Expression_of_ISGs_auto_sort, fig.width=12, fig.height=9}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
t(vsc_exprs)[,ISGs] %>%
    pheatmap(
      scale = 'column',
      fontsize = 8,
      show_rownames = FALSE,
      annotation_row = annotation_info,
      annotation_colors = group_pal,
      angle_col = 315,
      breaks = zscore_range,
      color = viridis(n = length(zscore_range)-1, option = "D"),
      cluster_rows = TRUE,
      border_color = NA)
```

\newpage

## Expression of individual genes in the M1.2, M3.4, and M5.12 gene modules {.tabset .tabset-fade .tabset-pills}

```{r Expression_of_individual_genes_in_the_M1.2, M3.4, and M5.12 gene modules, fig.height=6, fig.width=9}
tar_load(module_ISGs)
```

### Sorted by responder status

```{r module_ISGs_disease_sort, fig.width=12, fig.height=9}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
t(vsc_exprs)[sample_disease_order,(dplyr::arrange(as_tibble(module_ISGs, rownames = "gene"), module)$gene)] %>%
  pheatmap(
    scale = 'column',
    fontsize = 6,
    show_rownames = FALSE,
    annotation_row = annotation_info,
    annotation_col = module_ISGs,
    annotation_colors = group_pal,
    cluster_cols = FALSE,
    cluster_rows = FALSE,
    angle_col = 315,
    breaks = zscore_range,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    gaps_row = sample_disease_order_count,
    border_color = NA)
```

### Sorted by hierarchical clustering

```{r module_ISGs_auto_sort, fig.width=12, fig.height=9}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
t(vsc_exprs)[sample_disease_order,(dplyr::arrange(as_tibble(module_ISGs, rownames = "gene"), module)$gene)] %>%
  pheatmap(
    scale = 'column',
    fontsize = 6,
    show_rownames = FALSE,
    annotation_row = annotation_info,
    annotation_col = module_ISGs,
    annotation_colors = group_pal,
    cluster_cols = FALSE,
    cluster_rows = TRUE,
    angle_col = 315,
    breaks = zscore_range,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    border_color = NA
    )
```

\newpage

# Sample Clustering

The eigenvalues for the first 100 principal components were used to calculate a random forest-based adjacentcy matrix, which was in turn used to calculate the gap statistic to determine the optimal *k*-clusters.

```{r Sample_Clustering_chart}
annotation_info %>%
  dplyr::mutate(cluster = as_factor(cluster)) %>%
  group_by({{diffused_comparison}},
           cluster) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = cluster,
              values_from = n) %>%
  DT::datatable()

annotation_info %>%
  dplyr::mutate(cluster = as_factor(cluster)) %>%
  dplyr::select({{diffused_comparison}}, cluster) %>%
  group_by({{diffused_comparison}}) %>%
  dplyr::mutate(total_disease = n()) %>%
  dplyr::ungroup() %>%
  group_by({{diffused_comparison}}, cluster) %>%
  dplyr::mutate(total_cluster_disease = n(),
         freq = total_cluster_disease/total_disease,
         cluster = as_factor(cluster)) %>%
  distinct() %>%
  ggplot(aes(
    x = {{diffused_comparison}},
    y = freq,
    group = cluster,
    fill = cluster
  )) +
  geom_col() +
  labs(y = "Proportion of disease class in cluster") +
  scale_fill_manual(values = group_pal$cluster) +
  theme_pubr()
```

\newpage

# Clustering and variation of samples in reduced dimensional space {.tabset .tabset-fade .tabset-pills}

## PCA {.tabset .tabset-fade .tabset-pills}

```{r}
tar_load(pca_results)
```

\newpage

### PCA by disease classification

```{r PCA_by_groupification, fig.width=12, fig.height=6}
p1 <-
  ggplot(
    data = pca_results,
    aes(
      x = PC1,
      y = PC2,
      fill = {{diffused_comparison}},
      line = 1
    ),
    shape = 1
  ) +
    geom_point(
      shape = 21,
      colour = "black",
      size = 2.5,
      stroke = 0.5,
      alpha = 0.6
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme_pubr() +
    theme(legend.position = "none")

p2 <-
  ggplot(
    data = pca_results,
    aes(
      x = PC1,
      y = PC2,
      fill = {{diffused_comparison}},
      line = 1
      ), shape = 1
    ) +
  geom_point(
    shape = 21,
    colour = "black",
    size = 2.5,
    stroke = 0.5,
    alpha = 0.6
    ) +
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  theme_pubr() +
  geom_mark_ellipse(aes(fill = {{diffused_comparison}},
                        color = {{diffused_comparison}}),
                    alpha = 0.05
                    ) +
  theme(legend.position = "bottom")

legend <- get_legend(p2)

plot_grid(
  plot_grid(
    p1 + theme(legend.position = "none"),
    p2 + theme(legend.position = "none")
    ),
  legend,
  nrow = 2,
  rel_heights = c(1, .2)
)
```

\newpage

### PCA by plate

```{r PCA_by_plate, fig.width=12, fig.height=7}
p1 <- ggplot(
  data = pca_results,
  mapping = aes(
    x = PC1,
    y = PC2,
    fill = run_id
    )
  ) +
  geom_point(
    shape = 21,
    colour = "black",
    size = 2.5,
    stroke = 0.5,
    alpha = 0.6
    ) +
  labs(color = "run_id") +
  scale_color_paletteer_d("ggsci::default_igv") +
  scale_fill_paletteer_d("ggsci::default_igv")

p2 <- ggplot(
  data = pca_results,
  mapping = aes(
    x = PC1,
    y = PC2,
    fill = run_id
    )
  ) +
  geom_point(
    shape = 21,
    colour = "black",
    size = 2.5,
    stroke = 0.5,
    alpha = 0.6
    ) +
  labs(color = "run_id") +
  geom_mark_ellipse(aes(
    fill = run_id,
    color = run_id
    ),
    alpha = 0.05
    ) +
  scale_color_paletteer_d("ggsci::default_igv") +
  scale_fill_paletteer_d("ggsci::default_igv") +
  guides(color = guide_legend(nrow = 4)) +
  theme(legend.position = "bottom")

legend <- get_legend(p2)

plot_grid(
  plot_grid(
    p1 + theme(legend.position = "none"),
    p2 + theme(legend.position = "none")
    ),
  legend,
  nrow = 2,
  rel_heights = c(1, .2)
)
```

\newpage

### PCA by ethnicity

```{r PCA_by_ethnicity, fig.width=12, fig.height=6}
p1 <-
  pca_results %>%
  dplyr::select(
    PC1,
    PC2,
    sample_name,
    ethnicity
    ) %>%
  dplyr::mutate(ethnicity = as_factor(ethnicity)) %>%
  ggplot(
    aes(
      x = PC1,
      y = PC2,
      fill = ethnicity,
      line = 1
    ),
    shape = 1
  ) +
    geom_point(
      shape = 21,
      colour = "black",
      size = 2.5,
      stroke = 0.5,
      alpha = 0.6
    ) +
    scale_fill_paletteer_d(palette = "ggthemes::calc") +
    theme_pubr() +
    theme(legend.position = "none")

p2 <-
  pca_results %>%
  dplyr::select(PC1, PC2, sample_name, ethnicity) %>%
  dplyr::mutate(ethnicity = as_factor(ethnicity)) %>%
  ggplot(
    aes(
      x = PC1,
      y = PC2,
      fill = ethnicity,
      line = 1
      ), shape = 1
    ) +
  geom_point(
    shape = 21,
    colour = "black",
    size = 2.5,
    stroke = 0.5,
    alpha = 0.6
    ) +
  scale_fill_paletteer_d(palette = "ggthemes::calc") +
  scale_color_paletteer_d(palette = "ggthemes::calc") +
  theme_pubr() +
  geom_mark_ellipse(aes(fill = ethnicity,
                        color = ethnicity),
                    alpha = 0.05
                    ) +
  theme(legend.position = "bottom")

legend <- get_legend(p2)

plot_grid(
  plot_grid(
    p1 + theme(legend.position = "none"),
    p2 + theme(legend.position = "none")
    ),
  legend,
  nrow = 2,
  rel_heights = c(1, .2)
)
```

\newpage

<!-- ### 3D PCA colored by K-means cluster -->

<!-- ```{r 3D_PCA_colored_by_K-means cluster} -->

<!-- pca_results %>% -->

<!--   plot_ly( -->

<!--     x = ~PC1, -->

<!--     y = ~PC2, -->

<!--     z = ~PC3, -->

<!--     colors = group_pal$cluster -->

<!--   ) %>% -->

<!--   add_markers( -->

<!--     text = str_glue("Sample: {pca_results$sample_name}<br>Classification: {pca_results[[comparison_grouping_variable]]}<br>Ethnicity {pca_results$ethnicity}<br>Cluster: {pca_results$cluster}<br>Sex: {pca_results$sex}<br>Plate: {pca_results$run_id}"), -->

<!--     color = ~cluster, -->

<!--     size = 2.5, -->

<!--     alpha = 0.9, -->

<!--     hovertemplate = "%{text}", -->

<!--     marker = list( -->

<!--       size = 3, -->

<!--       line = list( -->

<!--         color = 'rgba(0, 0, 0, .8)', -->

<!--         width = 1.5 -->

<!--       ) -->

<!--     ) -->

<!--   ) -->

<!-- ``` -->

\newpage

## UMAP {.tabset .tabset-fade .tabset-pills}

```{r}
tar_load(umap_results)
```

### UMAP by disease classification

```{r UMAP_by_groupification, fig.width=12, fig.height=7}
p1 <-
  ggplot(
    data = umap_results,
    aes(
      x = umap_1,
      y = umap_2,
      fill = {{diffused_comparison}},
      line = 1
    ),
    shape = 1
  ) +
    geom_point(
      shape = 21,
      colour = "black",
      size = 2.5,
      stroke = 0.5,
      alpha = 0.6
    ) +
    scale_fill_brewer(palette = "Set1") +
    theme_pubr() +
    theme(legend.position = "none")

p2 <-
  ggplot(
    data = umap_results,
    aes(
      x = umap_1,
      y = umap_2,
      fill = {{diffused_comparison}},
      line = 1
    ), shape = 1
  ) +
    geom_point(
      shape = 21,
      colour = "black",
      size = 2.5,
      stroke = 0.5,
      alpha = 0.6
    ) +
    scale_fill_brewer(palette = "Set1") +
    scale_color_brewer(palette = "Set1") +
    theme_pubr() +
    geom_mark_ellipse(aes(fill = {{diffused_comparison}},
                          color = {{diffused_comparison}}),
                      alpha = 0.05
    ) +
  guides(color = guide_legend(nrow = 4)) +
  theme(legend.position = "bottom")

legend <- get_legend(p2)

plot_grid(
  plot_grid(
    p1 + theme(legend.position = "none"),
    p2 + theme(legend.position = "none")
    ),
  legend,
  nrow = 2,
  rel_heights = c(1, .2)
)
```

\newpage

### UMAP by plate

```{r UMAP_by_plate, fig.width=12, fig.height=7}
p1 <- ggplot(
  data = umap_results,
  mapping = aes(
    x = umap_1,
    y = umap_2,
    fill = run_id
  )
) +
  geom_point(
    shape = 21,
    colour = "black",
    size = 2.5,
    stroke = 0.5,
    alpha = 0.6
  ) +
  labs(color = "run_id") +
  scale_color_paletteer_d("ggsci::default_igv") +
  scale_fill_paletteer_d("ggsci::default_igv")

p2 <- ggplot(
  data = umap_results,
  mapping = aes(
    x = umap_1,
    y = umap_2,
    fill = run_id
    )
  ) +
  geom_point(
    shape = 21,
    colour = "black",
    size = 2.5,
    stroke = 0.5,
    alpha = 0.6
    ) +
  labs(color = "run_id") +
  geom_mark_ellipse(aes(
    fill = run_id,
    color = run_id
    ),
    alpha = 0.05
    ) +
  scale_color_paletteer_d("ggsci::default_igv") +
  scale_fill_paletteer_d("ggsci::default_igv") +
  guides(color = guide_legend(nrow = 4)) +
  theme(legend.position = "bottom")

legend <- get_legend(p2)

plot_grid(
  plot_grid(
    p1 + theme(legend.position = "none"),
    p2 + theme(legend.position = "none")
    ),
  legend,
  nrow = 2,
  rel_heights = c(1, .2)
)
```

\newpage

### UMAP by ethnicity

```{r UMAP_by_ethnicity, fig.width=12, fig.height=6}
p1 <-
  umap_results %>%
  dplyr::select(umap_1, umap_2, sample_name, ethnicity) %>%
  dplyr::mutate(ethnicity = as_factor(ethnicity)) %>%
  ggplot(
    aes(
      x = umap_1,
      y = umap_2,
      fill = ethnicity,
      line = 1
    ),
    shape = 1
  ) +
    geom_point(
      shape = 21,
      colour = "black",
      size = 2.5,
      stroke = 0.5,
      alpha = 0.6
    ) +
    scale_fill_paletteer_d(palette = "ggthemes::calc") +
    theme_pubr() +
    theme(legend.position = "none")

p2 <-
  umap_results %>%
  dplyr::select(umap_1, umap_2, sample_name, ethnicity) %>%
  dplyr::mutate(ethnicity = as_factor(ethnicity)) %>%
  ggplot(
    aes(
      x = umap_1,
      y = umap_2,
      fill = ethnicity,
      color = ethnicity,
      line = 1
      ), shape = 1
    ) +
  geom_point(
    shape = 21,
    colour = "black",
    size = 2.5,
    stroke = 0.5,
    alpha = 0.6
    ) +
  scale_fill_paletteer_d(palette = "ggthemes::calc") +
  scale_color_paletteer_d(palette = "ggthemes::calc") +
  theme_pubr() +
  geom_mark_ellipse(
    alpha = 0.05
    ) +
  theme(legend.position = "bottom")

legend <- get_legend(p2)

plot_grid(
  plot_grid(
    p1 + theme(legend.position = "none"),
    p2 + theme(legend.position = "none")
    ),
  legend,
  nrow = 2,
  rel_heights = c(1, .2)
)
```

\newpage

### UMAP by K-means cluster

```{r UMAP_by_K-means cluster, fig.height=6, fig.width=7}
umap_results %>%
  ggplot(
    aes(
      x = umap_1,
      y = umap_2,
      fill = cluster
      )
    ) +
  geom_point(
    shape = 21,
    colour = "black",
    size = 2.5,
    stroke = 0.5
    ) +
  scale_fill_manual(values = group_pal$cluster)
```

<!-- ### UMAP in 3D by K-means cluster -->

<!-- ```{r UMAP_in_3D_by_K-means cluster} -->

<!-- umap_results %>% -->

<!--   plot_ly( -->

<!--     x = ~umap_1, -->

<!--     y = ~umap_2, -->

<!--     z = ~umap_3, -->

<!--     colors = group_pal$cluster -->

<!--   ) %>% -->

<!--   add_markers( -->

<!--     text = str_glue("Sample: {umap_results[["sample_name"]]}<br>Classification: {umap_results[[comparison_grouping_variable]]}<br>Ethnicity {umap_results[["ethnicity"]]}<br>Cluster: {umap_results[["cluster"]]}<br>Sex: {umap_results[["sex"]]}<br>Plate: {umap_results[["run_id"]]}"), -->

<!--     color = ~cluster, -->

<!--     size = 2.5, -->

<!--     alpha = 0.9, -->

<!--     hovertemplate = "%{text}", -->

<!--     marker = list( -->

<!--       size = 3, -->

<!--       line = list( -->

<!--         color = 'rgba(0, 0, 0, .8)', -->

<!--         width = 1.5 -->

<!--       ) -->

<!--     ) -->

<!--   ) -->

<!-- ``` -->

\newpage

# Module scores {.tabset .tabset-fade .tabset-pills}

## SLE Interferon module scores {.tabset .tabset-fade .tabset-pills}

```{r SLE_Interferon_module_scores}
tar_load(module_scores_with_viral)
tar_load(ifn_modules)
```

### Sorted by disease classification

```{r SLE_Interferon_module_scores_disease_sort, fig.width=4, fig.height=6}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
module_scores_with_viral %>%
  dplyr::select(sample_name, one_of(ifn_modules)) %>%
  dplyr::arrange(match(sample_name, sample_disease_order)) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale="column",
    fontsize = 9,
    show_rownames = FALSE,
    annotation_colors = group_pal,
    annotation_row = annotation_info,
    breaks = zscore_range,
    cluster_rows = FALSE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    gaps_row = sample_disease_order_count,
    border_color = NA)
```

\newpage

### Sorted by cluster

```{r SLE_Interferon_module_scores_cluster_sort, fig.width=4, fig.height=6}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
module_scores_with_viral %>%
  dplyr::select(sample_name, one_of(ifn_modules)) %>%
  dplyr::arrange(match(sample_name, sample_cluster_order)) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale="column",
    fontsize = 9,
    show_rownames = FALSE,
    annotation_colors = group_pal,
    annotation_row = annotation_info,
    breaks = zscore_range,
    cluster_rows = FALSE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    gaps_row = sample_cluster_order_count,
    border_color = NA)
```

\newpage

### Sorted by hierarchical clustering

```{r SLE_Interferon_module_scores_auto_sort, fig.width=5, fig.height=6}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
module_scores_with_viral %>%
  dplyr::select(sample_name, one_of(ifn_modules)) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale="column",
    fontsize = 9,
    show_rownames = FALSE,
    annotation_colors = group_pal,
    annotation_row = annotation_info,
    breaks = zscore_range,
    cluster_rows = TRUE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    border_color = NA)
```

\newpage

## SLE Inflammatory module scores {.tabset .tabset-fade .tabset-pills}

```{r SLE_Inflammatory_module_scores}
tar_load(inflame_modules)
```

### Sorted by disease classification

```{r SLE_Inflammatory_module_scores_disease_sort, fig.width=4.5, fig.height=6}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
module_scores_with_viral %>%
  dplyr::select(sample_name, one_of(inflame_modules)) %>%
  dplyr::arrange(match(sample_name, sample_disease_order)) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale="column",
    fontsize = 9,
    show_rownames = FALSE,
    annotation_colors = group_pal,
    annotation_row = annotation_info,
    breaks = zscore_range,
    cluster_rows = FALSE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    gaps_row = sample_disease_order_count,
    border_color = NA)
```

\newpage

### Sorted by cluster

```{r SLE_Inflammatory_module_scores_cluster_sort, fig.width=4.5, fig.height=6}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
module_scores_with_viral %>%
  dplyr::select(sample_name, one_of(inflame_modules)) %>%
  dplyr::arrange(match(sample_name, sample_cluster_order)) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale="column",
    fontsize = 9,
    show_rownames = FALSE,
    annotation_colors = group_pal,
    annotation_row = annotation_info,
    breaks = zscore_range,
    cluster_rows = FALSE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    gaps_row = sample_cluster_order_count,
    border_color = NA)
```

\newpage

### Sorted by hierarchical clustering

```{r SLE_Inflammatory_module_scores_auto_sort, fig.width=5, fig.height=6}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
module_scores_with_viral %>%
  dplyr::select(sample_name, one_of(inflame_modules)) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale="column",
    fontsize = 9,
    show_rownames = FALSE,
    annotation_colors = group_pal,
    annotation_row = annotation_info,
    breaks = zscore_range,
    cluster_rows = TRUE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    border_color = NA)
```

## Low-density granulocyte module scores {.tabset .tabset-fade .tabset-pills}

```{r Low-density granulocyte module scores}
tar_load(ldg_modules)
range_bound = 3
zscore_range <- c(-range_bound:range_bound)
```

### Sorted by disease classification

```{r ldg_module_disease_sort, fig.width=4.5, fig.height=6}
module_scores_with_viral %>%
    dplyr::select(sample_name,
         one_of(names(ldg_modules))) %>%
  dplyr::arrange(match(sample_name, sample_disease_order)) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale="column",
    fontsize = 9,
    show_rownames = FALSE,
    annotation_colors = group_pal,
    annotation_row = annotation_info,
    breaks = zscore_range,
    cluster_rows = FALSE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    gaps_row = sample_disease_order_count,
    border_color = NA)
```

### Sorted by cluster

```{r ldg_module_cluster_sort, fig.width=4.5, fig.height=6}
module_scores_with_viral %>%
    dplyr::select(sample_name,
         one_of(names(ldg_modules))) %>%
  dplyr::arrange(match(sample_name, sample_cluster_order)) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale="column",
    fontsize = 9,
    show_rownames = FALSE,
    annotation_colors = group_pal,
    annotation_row = annotation_info,
    breaks = zscore_range,
    cluster_rows = FALSE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    gaps_row = sample_cluster_order_count,
    border_color = NA)
```

### Sorted by hierarchical clustering

```{r ldg_module_auto_sort, fig.width=5, fig.height=6}
module_scores_with_viral %>%
    dplyr::select(sample_name,
         one_of(names(ldg_modules))) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale="column",
    fontsize = 9,
    show_rownames = FALSE,
    annotation_colors = group_pal,
    annotation_row = annotation_info,
    breaks = zscore_range,
    cluster_rows = TRUE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    border_color = NA)
```

## SLE Modules with an annotated associated pathway {.tabset .tabset-fade .tabset-pills}

```{r}
tar_load(annotated_modules)
tar_load(annotated_mod_list)
```

```{r SLE_Modules_with_an_annotated_associated_pathway, fig.height=9, fig.width=11}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)

module_order <- c(
  "M1.1",
  
  "M2.3",
  "M3.1",
  "M6.18",
  
  "M1.2",
  "M3.4",
  "M5.12",
  
  "M3.2",
  "M4.2",
  "M4.6",
  "M4.13",
  "M5.1",
  "M5.7",
  "M7.1",
  
  "M3.6",
  "M8.46",
  "M4.1",
  "M4.15",
  "M4.10",
  "M4.11",
  
  "M4.14",
  "M5.15",
  "ldg_1.1",
  "ldg_2.1",
  "M8.83",
  
  "M6.6",
  "M2.2",
  "M3.3",
  "M3.5",
  "M4.7",
  "M6.11",
  "M6.16",
  "M9.42",
  "M6.13",
  "M4.3",
  "M4.5",
  "M5.9",
  "M5.6",
  "M5.10",
  "M6.2",
  "M6.12"
)

module_gaps <- c(1,4,7,14,20,25)
```

### Sorted by responder status

```{r annotated_module_disease_sort, fig.width=12, fig.height=9}
module_scores_with_viral %>%
  dplyr::arrange({{diffused_comparison}}) %>%
  dplyr::select(
    sample_name,
    one_of(module_order)
  ) %>%
  pivot_longer(
    cols = one_of(module_order),
    names_to = "module",
    values_to = "score"
    ) %>%
  dplyr::mutate(
    module =
      recode(
        module,
        !!!annotated_mod_list
        )
    ) %>%
  pivot_wider(
    names_from = "module",
    values_from = "score"
    ) %>%
  column_to_rownames("sample_name") %>%
  t() %>%
  pheatmap(
    scale="row",
    fontsize = 9,
    show_rownames = TRUE,
    show_colnames = FALSE,
    annotation_colors = group_pal,
    annotation_row = column_to_rownames(dplyr::mutate(annotated_modules, module = recode(module, !!!annotated_mod_list)), var = "module"),
    annotation_col =
      dplyr::arrange(
        .data = annotation_info[,comparison_grouping_variable,drop=FALSE],
        {{diffused_comparison}}
        ),
    breaks = zscore_range,
    cluster_rows = FALSE,
    cluster_cols = FALSE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    gaps_col = sample_disease_order_count,
    gaps_row = module_gaps,
    border_color = NA)
```

### Sorted by hierarchical clustering

```{r annotated_module_auto_sort, fig.width=12, fig.height=9}
module_scores_with_viral %>%
  dplyr::select(sample_name,
         one_of(module_order)) %>%
  column_to_rownames("sample_name") %>%
  t() %>%
  pheatmap(
    scale="row",
    fontsize = 9,
    show_rownames = FALSE,
    show_colnames = TRUE,
    annotation_colors = group_pal,
    annotation_row = column_to_rownames(annotated_modules, var = "module"),
    annotation_col = annotation_info[,c("cluster", comparison_grouping_variable),drop=FALSE],
    breaks = zscore_range,
    cluster_rows = FALSE,
    cluster_cols = TRUE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    angle = 315,
    gaps_row = module_gaps,
    border_color = NA)
```

\newpage

## All SLE modules {.tabset .tabset-fade .tabset-pills}

```{r All_SLE_modules, fig.height=9, fig.width=11}
tar_load(module_annotation)
range_bound = 3
zscore_range <- c(-range_bound:range_bound)
```

### Sorted by responder status

```{r all_module_disease_sort, fig.width=12, fig.height=9}
module_scores_with_viral %>%
  dplyr::select(
    sample_name,
    matches("^M[[:digit:]]+")
    ) %>%
  dplyr::arrange(
    match(
      sample_name,
      sample_disease_order
      )
    ) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale = 'column',
    fontsize = 9,
    annotation_row = annotation_info,
    annotation_col = column_to_rownames(module_annotation, var = "module"),
    annotation_colors = group_pal,
    show_rownames = FALSE,
    show_colnames = FALSE,
    angle_col = 315,
    breaks = zscore_range,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    gaps_row = sample_disease_order_count,
    border_color = NA,
    cluster_rows = FALSE
  )
```

### Sorted by hierarchical clustering

```{r all_module_auto_sort, fig.width=12, fig.height=9}
module_scores_with_viral %>%
  dplyr::select(
    sample_name,
    matches("^M[[:digit:]]+")
    ) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale = 'column',
    fontsize = 9,
    annotation_row = annotation_info,
    annotation_col = column_to_rownames(module_annotation, var = "module"),
    annotation_colors = group_pal,
    show_rownames = FALSE,
    show_colnames = FALSE,
    angle_col = 315,
    breaks = zscore_range,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    border_color = NA
  )
```

\newpage

## Expression of the annotated Banchereau modules {.tabset .tabset-fade .tabset-pills}

```{r}
tar_load(module_scores_with_viral)
tar_load(annotated_module_scores)
tar_load(annotated_module_scores_pivot)
tar_load(annotated_module_stats_by_disease)

number_of_modules <-
  ncol(annotated_module_scores) - 1

number_of_pages <-
  ceiling(number_of_modules/9)
```

### By responder status

```{r Banchereau_modules_by_disease, echo=FALSE, fig.height=8, fig.width=12, warning=FALSE}
annotated_module_group_graphs <-
  annotated_module_scores_pivot %>%
  dplyr::select(-cluster) %>%
  dplyr::mutate({{diffused_comparison}} := as_factor({{diffused_comparison}})) %>%
  ggviolin(
    x = comparison_grouping_variable,
    y = "score",
    fill = comparison_grouping_variable,
    draw_quantiles = c(0.25, 0.5, 0.75),
    add = "jitter",
    trim = TRUE,
    add.params = list(size = 1, alpha = 0.25)
    ) +
  scale_fill_paletteer_d("ggsci::uniform_startrek") +
  stat_pvalue_manual(
    data = annotated_module_stats_by_disease,
    label = 
      ifelse(
        test = ("p.adj.signif" %in% annotated_module_stats_by_disease),
        yes = "p.adj.signif",
        no = "p"
      ),
    hide.ns = TRUE,
    bracket.size = 0.1,
    tip.length = 0.01,
    size = 4
  ) +
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 9),
        axis.text.y = element_text(size = 9))
  
for(i in seq(number_of_pages)){
  print(
    annotated_module_group_graphs +
      facet_wrap_paginate(
        facets = vars(module),
        scales = "free_y",
        nrow = 2,
        ncol = 3,
        page = i)
  )
}
```

### By cluster

```{r Banchereau_modules_by_cluster, echo=FALSE, fig.height=8, fig.width=12, warning=FALSE}
number_of_modules <-
  ncol(annotated_module_scores) - 1

number_of_pages <-
  ceiling(number_of_modules/9)
  
annotated_module_cluster_graphs <-
  annotated_module_scores_pivot %>%
  dplyr::select(-{{diffused_comparison}}) %>%
  dplyr::mutate(cluster = as_factor(cluster)) %>%
  ggviolin(
    x = "cluster",
    y = "score",
    fill = "cluster",
    draw_quantiles = c(0.25, 0.5, 0.75),
    add = "jitter",
    trim = TRUE,
    add.params = list(size = 1, alpha = 0.5)
    ) +
  scale_fill_paletteer_d("ggsci::uniform_startrek") +
  stat_pvalue_manual(
    data = annotated_module_stats_by_cluster,
    label = 
      ifelse(
        test = ("p.adj.signif" %in% annotated_module_stats_by_cluster),
        yes = "p.adj.signif",
        no = "p"
      ),
    hide.ns = TRUE,
    bracket.size = 0.1,
    tip.length = 0.01,
    size = 4
  ) +
  theme_pubr() +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1,
        vjust = 1,
        size = 9
        ),
    axis.text.y =
      element_text(size = 9)
    )
  

for(i in seq(number_of_pages)){
  print(
    annotated_module_cluster_graphs +
      facet_wrap_paginate(
        facets = vars(module),
        scales = "free_y",
        nrow = 2,
        ncol = 3,
        page = i)
  )
}
```

\newpage

# WGCNA

```{r include=FALSE}
tar_load(wgcna_modules)
```

## Expression similarity dendrogram

```{r WGCNA_Expression_similarity_dendrogram, echo=FALSE, fig.height=6, fig.width=10}
plotDendroAndColors(
  dendro = wgcna_modules$dendrograms[[1]],
  colors = wgcna_modules$colors[wgcna_modules$blockGenes[[1]]],
  groupLabels = "Module colors",
  main = "Gene dendrogram and module colors in block 1",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05
)
```

## Expression of identified modules {.tabset .tabset-fade .tabset-pills}

```{r include=FALSE}
tar_load(module_scores_with_viral_by_disease)
tar_load(module_scores_with_viral_by_disease_stats)

mods <- module_scores_with_viral_by_disease %>% dplyr::pull(module) %>% unique()
module_bins <- split(mods, ceiling(seq_along(mods)/9))
```

### By responder status

```{r wgcna_by_disease, echo=FALSE, fig.height=8, fig.width=12}
module_disease_graphs <-
  module_scores_with_viral_by_disease %>%
  ggviolin(
    x = comparison_grouping_variable,
    y = "score",
    fill = comparison_grouping_variable,
    draw_quantiles = c(0.25, 0.5, 0.75),
    add = "jitter",
    trim = TRUE,
    add.params = list(size = 1, alpha = 0.25)
    ) +
  scale_fill_paletteer_d("ggsci::uniform_startrek") +
  stat_pvalue_manual(
    data = module_scores_with_viral_by_disease_stats,
    label =
      ifelse(
        test = ("p.adj.signif" %in% module_scores_with_viral_by_disease_stats),
        yes = "p.adj.signif",
        no = "p"
        ),
    hide.ns = TRUE,
    bracket.size = 0.1,
    tip.length = 0.01,
    size = 4
  ) +
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 9),
        axis.text.y = element_text(size = 9),
        plot.caption = element_markdown()) +
  labs(caption = "Mann-Whitney, Benjamini–Hochberg adjusted value.\n *<i>p</i> <0.05, **<i>p</i> <0.01, ***<i>p</i> <0.001, ****<i>p</i> <0.0001")

for(i in seq(module_bins)){

  g <-
    module_scores_with_viral_by_disease %>%
    dplyr::filter(module %in% module_bins[[i]]) %>%
    ggviolin(
      x = "disease_class",
      y = "score",
      fill = "disease_class",
      draw_quantiles = c(0.25, 0.5, 0.75),
      add = "jitter",
      trim = TRUE,
      add.params = list(size = 1, alpha = 0.25)
      ) +
    scale_fill_paletteer_d("ggsci::uniform_startrek") +
    stat_pvalue_manual(
      data = dplyr::filter(module_scores_with_viral_by_disease_stats, module %in% module_bins[[i]]),
      label = "p",
      hide.ns = TRUE,
      bracket.size = 0.1,
      tip.length = 0.01,
      size = 4
    ) +
    theme_pubr() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 9),
          axis.text.y = element_text(size = 9),
          plot.caption = element_markdown()) +
    labs(caption = "Mann-Whitney, Benjamini–Hochberg adjusted value.\n *<i>p</i> <0.05, **<i>p</i> <0.01, ***<i>p</i> <0.001, ****<i>p</i> <0.0001") +
    facet_wrap(vars(module), scales = "free_y", ncol = 3, nrow = 3)
  
  print(g)
}
```

\newpage

### By cluster

```{r wgcna_modules_by_cluster, echo=FALSE, fig.height=8, fig.width=12}
for (i in seq(module_bins)){
  g <-
    module_scores_with_viral_by_cluster %>%
    dplyr::filter(module %in% module_bins[[i]]) %>%
    ggviolin(
      x = "cluster",
      y = "score",
      fill = "cluster",
      draw_quantiles = c(0.25, 0.5, 0.75),
      add = "jitter",
      trim = TRUE,
      add.params = list(size = 1, alpha = 0.25)
      ) +
    scale_fill_paletteer_d("ggsci::uniform_startrek") +
    stat_pvalue_manual(
      data = dplyr::filter(module_scores_with_viral_by_cluster_stats, module %in% module_bins[[i]]),
      label = "p",
      hide.ns = TRUE,
      bracket.size = 0.1,
      tip.length = 0.01,
      size = 4
    ) +
  guides(fill = guide_legend(title = NULL)) +
  labs(caption = "Mann-Whitney, Benjamini–Hochberg adjusted value.\n *<i>p</i> <0.05, **<i>p</i> <0.01, ***<i>p</i> <0.001, ****<i>p</i> <0.0001")

for(i in seq(number_of_pages)){
  print(
    module_disease_graphs +
      facet_wrap_paginate(
        facets = vars(module),
        scales = "free_y",
        nrow = 3,
        ncol = 3,
        page = i
        )
  )
}
```

\newpage

## WGCNA eigenvalues {.tabset .tabset-fade .tabset-pills}

```{r wgcna_eigenvalues, fig.height=9, fig.width=11}
range_bound = 4
zscore_range <- c(-range_bound:range_bound)
```

### Sorted by responder status

```{r WGCNA_eigenvalues_disease_sort, fig.width=12, fig.height=9}
module_scores_with_viral %>%
  dplyr::select(
    sample_name,
    matches("^ME")
    ) %>%
  dplyr::arrange(
    match(
      sample_name,
      sample_disease_order
      )
    ) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale = 'column',
    fontsize = 9,
    annotation_row = annotation_info,
    annotation_colors = group_pal,
    show_rownames = FALSE,
    show_colnames = TRUE,
    angle_col = 315,
    breaks = zscore_range,
    cluster_rows = FALSE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    gaps_row = sample_disease_order_count,
    border_color = NA
    )
```

### Sorted by hierarchical clustering

```{r WGCNA_eigenvalues_auto_sort, fig.width=12, fig.height=9}
module_scores_with_viral %>%
  dplyr::select(
    sample_name,
    matches("^ME")
    ) %>%
  column_to_rownames("sample_name") %>%
  pheatmap(
    scale = 'column',
    fontsize = 9,
    annotation_row = annotation_info,
    annotation_colors = group_pal,
    show_rownames = FALSE,
    show_colnames = TRUE,
    angle_col = 315,
    breaks = zscore_range,
    cluster_rows = TRUE,
    color = viridis(n = length(zscore_range)-1, option = "D"),
    border_color = NA
    )
```

\newpage

## Gene ontology GSEA of WGCNA modules

```{r wgcna_gsea, fig.height=11, fig.width=12}
tar_load(MEplotting)

modules <- unique(MEplotting$module)
n <- 6

bins <-
  ceiling(
    x = seq_along(along.with = modules)/n
    ) %>%
  set_names(nm = modules) %>%
  enframe()

for (i in unique(bins[['value']])){
  modules_to_plot <-
    bins %>%
    dplyr::filter(
      value == i
    ) %>%
    dplyr::pull(name)
  
  module_data <-
    dplyr::filter(
      .data = MEplotting,
      module %in% modules_to_plot
    )
  
  plots <-
    module_data %>%
    ggplot(
      mapping =
        aes(
          y = GeneRatio,
          x = order,
          color = p.adjust
        )
    ) +
    geom_point(
      stat = "identity",
      size = 2
    ) +
    scale_x_continuous(
      breaks = module_data[["order"]],
      labels =
        str_wrap(
          module_data[["ID"]],
          width = 40
        ),
      expand = c(0.1,0.1)
    ) +
    scale_color_gradient(
      low = "blue",
      high = "red"
    ) +
    labs(color = "adjusted\np value") +
    facet_wrap(
      facets = vars(module),
      ncol = 2,
      nrow = 3,
      scales = "free",
    ) +
    theme(
      axis.text.y = element_text(size = 9),
      axis.text.x = element_text(size = 9),
      panel.background =
        element_rect(
          fill = "white",
          linetype = "solid"
        ),
      panel.border =
        element_rect(
          colour = "black",
          fill=NA, 
          size=1
        ),
      panel.grid = element_line(color = "grey90"),
      axis.title.y = element_blank()) +
    coord_flip()
  print(plots)
}
```

\newpage

## Use of WGCNA modules in classifying clusters.

## Use of WGCNA modules in labeling responder status.

Note: Subjects with LP were excluded due to the small population size.

```{r wgcna_cluster_confusion}
tar_load(wgcna_cluster_rf_cv)
tar_load(wgcna_cluster_split)

caret::confusionMatrix(
  data = predict(wgcna_cluster_rf_cv, testing(wgcna_cluster_split)),
  reference = as_factor(testing(wgcna_cluster_split)$cluster)
)
```

The relative importance of each eigengene in classification:

```{r wgcna_cluster_importance}
tar_load(wgcna_cluster_rf_cv_varImp)

wgcna_cluster_rf_cv_varImp
```

\newpage

## Use of WGCNA modules in labeling disease class.

The module eigengene scores were used to train a random forest model with repeated cross validation to predict cluster identity.

The model was trained with 75% of the dataset and tested on 25%.

```{r wgcna_disease_confusion}
tar_load(wgcna_group_rf_cv)
tar_load(wgcna_group_test)

caret::confusionMatrix(
  data = predict(wgcna_group_rf_cv, wgcna_group_test),
  reference = as_factor(wgcna_group_test[[comparison_grouping_variable]])
)
```

The relative importance of each eigengene in classification:

```{r wgcna_disease_importance}
tar_read(wgcna_group_rf_cv_varImp)
```

\newpage

## Use of module modules in classifying clusters.

The module eigengene scores were used to train a random forest model with repeated cross validation to predict cluster identity.

The model was trained with 75% of the dataset and tested on 25%.

```{r Banchereau_cluster_confusion}
tar_load(module_cluster_rf_cv)
tar_load(module_cluster_test)

caret::confusionMatrix(
  data = predict(module_cluster_rf_cv, module_cluster_test),
  reference = as_factor(module_cluster_test[["cluster"]])
)
```

The relative importance of each eigengene in classification:

```{r Banchereau_cluster_importance}
tar_read(module_cluster_rf_cv_varImp)
```

\newpage

## Use of module modules in labeling disease class.

The module eigengene scores were used to train a random forest model with repeated cross validation to predict responder status.

The model was trained with 75% of the dataset and tested on 25%.

```{r Banchereau_disease_confusion}
tar_load(module_group_rf_cv)
tar_load(module_group_test)

caret::confusionMatrix(
  data = predict(module_group_rf_cv, module_group_test),
  reference = as_factor(module_group_test[[comparison_grouping_variable]])
)
```

The relative importance of each eigengene in classification:

```{r Banchereau_disease_importance}
tar_read(module_group_rf_cv_varImp)
```

```{r}
sessioninfo::session_info()
```
