---
params: 
  set_title:
    label: "report title"
    value: "Project"
    # input: text
  set_author: 
    label: "author"
    value: "Miles Smith"
    # input: text
title: "`r params$set_title`"
author: "`r params$set_author`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    highlight: pygments
    toc: yes
    toc_float: yes
    keep_md: yes
  pdf_document:
    highlight: "pygments"
    toc: TRUE
    toc_depth: 3
always_allow_html: true
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 1500px !important;
      width: 1500px !important;
    }
    body {
      max-width: 1500px !important;
    }
    caption {
      color: black;
      font-weight: bold;
      font-size: 1.0em;
    }
```

```{r setup, include=FALSE}
options(width = 1200)
knitr::opts_chunk$set(echo = FALSE, root.dir = here::here(), warning = FALSE, message=FALSE)
ggplot2::theme_set(ggpubr::theme_pubr())
comparison_grouping_variable <-
  project_params[["comparison_grouping_variable"]]
diffused_comparison <- rlang::sym(comparison_grouping_variable)
diffused_comparison <- rlang::enquo(diffused_comparison)
source("project_parameters.R")
source("code/14_report_funcs.R")
```

```{r}
targets::tar_load(study_md)
targets::tar_load(res)

if (!is.null(project_params[["heatmap_row_annotations"]])){
  heatmap_row_annotations <- list(project_params[["heatmap_row_annotations"]])
} else{
  heatmap_row_annotations <- list(colnames(dplyr::select(study_md, where(is.factor))))
}
```

```{r study_group_table}
if (knitr::is_html_output()) {
  study_md |>
    janitor::tabyl(comparison_grouping_variable) |>
    janitor::adorn_pct_formatting(digits = 2) |>
    knitr::kable() |>
    kableExtra::kable_styling(
      bootstrap_options =
        c(
          "striped",
          "hover",
          "condensed",
          "responsive"
          ),
      full_width = FALSE
      ) |>
    kableExtra::column_spec(
      1,
      bold = TRUE,
      border_right = TRUE,
      width = "10em"
      ) |>
    kableExtra::column_spec(
      2,
      border_right = TRUE,
      width = "10em"
      ) |>
    kableExtra::column_spec(
      3,
      border_right = FALSE,
      width = "10em"
      )
} else if(knitr::is_latex_output()){
  study_md |>
    janitor::tabyl(comparison_grouping_variable) |>
    dplyr::rename(
      `Disease classification` = comparison_grouping_variable,
      Number = n,
      Percent = percent
      ) |>
    flextable::qflextable() |>
    flextable::set_formatter(
      Percent = function(x) sprintf("%.01f", x*100)
    )
}
```

\newpage

# Differential Gene Expression {.tabset .tabset-fade .tabset-pills}

```{r DE_table}
results_tbl <-
  res |>
    # purrr::map(alt_summary, lfcThreshold = params[["lfcThreshold"]]) %>%
  purrr::map(alt_summary, lfcThreshold = 0.1) |>
    tibble::enframe() |>
    tidyr::unnest(cols = c(value)) |>
    dplyr::mutate(
      `up %` = round(
        100*(as.integer(up)/as.integer(total)),
        digits = 2
        ),
      `down %` = round(
        100*(as.integer(down)/as.integer(total)),
        digits = 2
        ),
      `low counts %` = round(
        100*(as.integer(lowcounts)/as.integer(total)),
        digits = 2
        ),
      name = stringr::str_replace_all(
        string = name,
        pattern = "_",
        replacement = " "
        )
    )

if (knitr::is_html_output()) {
  dplyr::select(
    .data = results_tbl,
    comparison = name,
    up,
    `up %`,
    down,
    `down %`,
    `mean count cutoff` = ft,
    `low counts` = lowcounts,
    `low counts %`
  ) %>%
  knitr::kable() |>
  kableExtra::kable_styling(
    bootstrap_options =
      c(
        "striped",
        "hover",
        "condensed",
        "responsive"
        )
    ) %>%
  kableExtra::footnote(
    paste(
      alt_summary(res[[1]])$total,
      "with nonzero total read count", "\nOutliers: ",
      alt_summary(res[[1]])$outlier)
    )
} else if (knitr::is_latex_output()){
  dplyr::select(
    .data = results_tbl,
    Comparison = name,
    `# Up\nregulated` = up,
    `% Up\nregulated` = `up %`,
    `# Down\nregulated` = down,
    `% Down\nregulated` = `down %`,
    `Mean count\ncutoff` = ft,
    `# Low\ncounts` = lowcounts,
    `% Low\ncounts` = `low counts %`
    ) |>
  flextable::qflextable() |>
  flextable::set_formatter(
    `% Up\nregulated` = \(x) sprintf("%.02f", x),
    `% Down\nregulated` = \(x) sprintf("%.02f", x),
    `% Low\ncounts` = \(x) sprintf("%.02f", x)
  ) |>
  flextable::bold(part = "header") |>
  flextable::align(
    align = "left",
    part = "body",
    j = 1
    )
  }
```

\newpage

## 25 genes with largest negative log-fold change in experimental group {.tabset .tabset-fade .tabset-pills}
```{r 25_top_down_DE, echo = FALSE}
targets::tar_load(down_tables)
```

```{r 25_top_down_DE_tbl, echo = FALSE, results='asis'}
if (knitr::is_html_output()){
  purrr::iwalk(
    .x = down_tables,
    .f = \(dtbl, dtbl_name)
    process_deg_kable(
        deg_table = dtbl,
        deg_table_name = dtbl_name,
        .color = "tomato",
        .direction = "down"
      ) |>
      print()
  )
} else if (knitr::is_latex_output()){
  purrr::iwalk(
    .x = down_tables,
    .f = \(dtbl, dtbl_name)
      process_deg_flextable(dtbl, dtbl_name) %>%
      knitr::knit_print() %>%
      cat()
  )
}
```

## 25 genes with largest positive log-fold change in experimental group {.tabset .tabset-fade .tabset-pills}
```{r 25_top_up_DE, echo = FALSE}
targets::tar_load(up_tables)
```

```{r 25_top_up_DE_tbl, echo = FALSE, results='asis'}
if (knitr::is_html_output()){
  purrr::iwalk(
    .x = up_tables,
    .f = \(utbl, utbl_name)
      process_deg_kable(
        deg_table = utbl,
        deg_table_name = utbl_name,
        .color = "lightgreen",
        .direction = "up"
      ) |>
      print()
  )
} else if (knitr::is_latex_output()){
  purrr::iwalk(
    .x = up_tables,
    .f = \(utbl, utbl_name)
      process_deg_flextable(utbl, utbl_name) |>
      knitr::knit_print() |>
      cat()
  )
}
```

\newpage
```{r}
targets::tar_load(deg_class)
targets::tar_load(vsc_exprs)
targets::tar_load(annotation_info)
targets::tar_load(group_pal)
targets::tar_load(up_tables)
targets::tar_load(down_tables)

comparison_groups <- project_params[["comparison_groups"]]
deg_comparisons <- unique(c(names(up_tables), names(down_tables)))

groupings <- list(
  project_params[["comparison_grouping_variable"]],
  "cluster",
  "hierarchical_clustering"
)

groupings <- rlang::set_names(x = groupings, nm = groupings)
```

<!-- TODO: This is kind of dumb for the moment.  It only really does one group comparison, that is we only calculate -->
<!-- say the comparisons for each experimental group within Disease Class, not within Disease Class and within cluster and within race -->
<!-- but a lot more will have to change before we can do anything with that here -->
<!-- TODO: Add GSEA of top DEGs -->
```{r top_class_DE_disease_sort, fig.width=12, fig.height=9}
if(length(up_tables) > 0 & length(down_tables) > 0 ){
  deg_heatmaps <-
    purrr::map(deg_comparisons, \(comparison_group){
      header <-
        knitr::knit_expand(
          text = "\n## Expression of top DE genes for {{GROUPING_VAR}}: {{COMPARISON_GROUPING}} {.tabset .tabset-fade .tabset-pills}",
          COMPARISON_GROUPING = comparison_group,
          GROUPING_VAR        = project_params[["comparison_grouping_variable"]]
        )
      each_comparison <-
        purrr::map(groupings, \(group){
          subheader <-
            knitr::knit_expand(
              text = "\n### Sorted by {{SORTING}} \n",
              SORTING = group
            )
          dgehm <-
            knitr::knit_expand(
              file = here::here("analysis/templates/comparisonHeatmaps.Rmd"),
              COMPARISON     = comparison_group,
              UPTABLE        = paste0("up_tables[['", comparison_group, "']]"),
              DOWNTABLE      = paste0("down_tables[['", comparison_group, "']]"),
              DATASET        = "vsc_exprs",
              METADATA       = "study_md",
              COMPARISON_GROUP = project_params[["comparison_grouping_variable"]],
              GROUPING       = group,
              PALETTES       = "group_pal",
              ROW_ANNOTATION = heatmap_row_annotations,
              FIG_WIDTH      = 12,
              FIG_HEIGHT     = 9
            )
          c(subheader, unlist(dgehm))
        })
      c(header, unlist(each_comparison))
    })
  template_res <- knitr::knit_child(text = unlist(deg_heatmaps), quiet = TRUE)
  cat(template_res, sep = '\n')
}
```

\newpage

```{r deg_enrichment, echo = FALSE}
targets::tar_load(up_enrichment)
targets::tar_load(up_enrichment_degs)
targets::tar_load(down_enrichment)
targets::tar_load(down_enrichment_degs)
```

```{r, results = 'asis'}
if (!is_null_list(up_enrichment_degs)){
  up_deg_enrichment <-
    purrr::map(deg_comparisons, \(comparison_group){
      header <-
        knitr::knit_expand(
          text = "\n## Pathway enrichment upgregulated genes for {{GROUPING_VAR}}: {{COMPARISON_GROUPING}} {.tabset .tabset-fade .tabset-pills}",
          COMPARISON_GROUPING = comparison_group,
          GROUPING_VAR        = project_params[["comparison_grouping_variable"]]
        )
      ontologies <-
        purrr::map(
          .x = names(up_enrichment[[comparison_group]]),
          .f = \(x){
            knitr::knit_expand(
              file = here::here("analysis/templates/degPathwayPlots.Rmd"),
              TITLE = x,
              ENRICH = paste0("up_enrichment[['", comparison_group, "']][['", x, "']]"),
              LFC = paste0("up_enrichment_degs[['", comparison_group, "']][['", x, "']]")
              )
            })
      c(header, unlist(ontologies))
    })
  template_res <- knitr::knit_child(text = unlist(up_deg_enrichment), quiet = TRUE)
  cat(template_res, sep = '\n')
}
```

\newpage

```{r, results = 'asis'}
if(!is_null_list(down_enrichment_degs)){
  down_deg_enrichment <-
    purrr::map(deg_comparisons, \(comparison_group){
      header <-
        knitr::knit_expand(
          text = "\n## Pathway enrichment downgregulated genes for {{GROUPING_VAR}}: {{COMPARISON_GROUPING}} {.tabset .tabset-fade .tabset-pills}",
          COMPARISON_GROUPING = comparison_group,
          GROUPING_VAR        = project_params[["comparison_grouping_variable"]]
        )
      ontologies <-
        purrr::map(
          .x = names(down_enrichment[[comparison_group]]),
          .f = \(x){
            knitr::knit_expand(
              file = here::here("analysis/templates/degPathwayPlots.Rmd"),
              TITLE = x,
              ENRICH = paste0("down_enrichment[['", comparison_group, "']][['", x, "']]"),
              LFC = paste0("down_enrichment_degs[['", comparison_group, "']][['", x, "']]")
              )
            })
      c(header, unlist(ontologies))
    })
  template_res <- knitr::knit_child(text = unlist(down_deg_enrichment), quiet = TRUE)
  cat(template_res, sep = '\n')
}
```

\newpage

## Volcano plot of gene expression {.tabset .tabset-fade .tabset-pills}

```{r Volcano_plot_of_gene_expression, results = 'asis'}
volcanoes <- purrr::map(
  .x = names(res),
  .f = \(x){
  header <-
      knitr::knit_expand(
        text       = "\n\n### Comparison of {{COMPARISON}}\n",
        COMPARISON = stringr::str_replace(string = x, pattern = "_", replacement = " ")
      )
  volcano_plot <-
    knitr::knit_expand(
      file       = here::here("analysis/templates/volcanoPlots.Rmd"),
      X          = paste0("res[['",x,"']]"),
      Y          = x,
      FIG_HEIGHT = 9,
      FIG_WIDTH  = 9
    )
    c(header, volcano_plot)
  })
template_res <- knitr::knit_child(text = unlist(volcanoes), quiet = TRUE)
cat(template_res, sep = '\n')
```

\newpage

```{r Expression_of_selected_interferon-stimulated genes, fig.height=6, fig.width=9}
targets::tar_load(ISGs)
targets::tar_load(module_ISGs)

IFN_module_genes <- rownames(module_ISGs)

module_groups = 
  list(
    "selected interferon-stimulated genes" = ISGs,
    "individual genes in the M1.2, M3.4, and M5.12 gene modules" = IFN_module_genes
)

row_groupings  <-
  list(
    project_params[["comparison_grouping_variable"]],
    "cluster",
    "hierarchical clustering"
    )

row_groupings <- rlang::set_names(x = row_groupings, nm = row_groupings)
group_pal2 <- c(group_pal, generatePalettes(module_ISGs, "module"))
```

```{r, echo=FALSE, results='asis'}
groups <-
  purrr::imap(module_groups, \(module, title){
    header <-
      knitr::knit_expand(
        text     = "\n\n## Expression of {{MODULE_GROUP}} {.tabset .tabset-fade .tabset-pills}\n",
        MODULE_GROUP = title
      )
    each_comparison <-
      purrr::map(row_groupings, \(group){
        subheader <-
          knitr::knit_expand(
            text    = "\n\n### Sorted by {{SORTING}}\n\n",
            SORTING = group
          )
        heatmap <-
          knitr::knit_expand(
            file           = here::here("analysis/templates/groupedComplexHeatmaps.Rmd"),
            DATASET        = "vsc_exprs",
            VARIABLE_LIST  = list(module),
            METADATA       = "study_md",
            PALETTES       = "group_pal2",
            ROW_GROUPING   = group,
            ROW_ANNOTATION = heatmap_row_annotations,
            COL_GROUPING   = "none",
            COL_ANNOTATION = 'tibble::as_tibble(module_ISGs, rownames="obsv")',
            FIG_WIDTH       = min(((1*length(module)) + 3.84), 12),
            FIG_HEIGHT      = 6,
            SCALE          = "TRUE"
          )
        c(subheader, heatmap)
      })
    c(header, unlist(each_comparison))
  })
template_res <- knitr::knit_child(text = unlist(groups), quiet = TRUE)
cat(template_res, sep = '\n')
```

\newpage

# Sample Clustering

The eigenvalues for the first 100 principal components were used to calculate a random forest-based adjacency matrix, which was in turn used to calculate the gap statistic to determine the optimal *k*-clusters.

```{r Sample_Clustering_chart, fig.height=6, fig.width=6}
annotation_info |>
  dplyr::count(
    {{diffused_comparison}},
    cluster
    ) |>
  tidyr::pivot_wider(
    names_from = cluster,
    values_from = n
    ) |>
  DT::datatable()

annotation_info |>
  tibble::as_tibble() |>
  dplyr::select(
    {{diffused_comparison}},
    cluster
  ) |>
  dplyr::add_count({{diffused_comparison}}, name = "total") |>
  dplyr::add_count({{diffused_comparison}}, cluster, name = "total_cluster") |>
  dplyr::mutate(freq = total_cluster/total) |>
  dplyr::distinct() |>
  ggplot2::ggplot(
    ggplot2::aes(
      x = {{diffused_comparison}},
      y = freq,
      group = cluster,
      fill = cluster
      )
    ) +
  ggplot2::geom_col() +
  ggplot2::labs(y = glue::glue("Proportion of {project_params[['comparison_grouping_variable']]} in cluster")) +
  ggplot2::scale_fill_manual(values = group_pal$cluster) +
  ggpubr::theme_pubr()
```

\newpage

# Clustering and variation of samples in reduced dimensional space {.tabset .tabset-fade .tabset-pills}

```{r setup_reductions, echo = FALSE}
targets::tar_load(pca_results)
targets::tar_load(umap_results)

reductions = list(
  PCA  = "pca",
  UMAP = "umap"
)

groupings = list(
  project_params[["comparison_grouping_variable"]],
  "cluster"
) |>
  rlang::set_names(
    nm = c(
      project_params[["comparison_grouping_variable"]],
      "K-means cluster"
      )
    )
```

```{r dimplots, results = "asis"}
dimplots <-
  purrr::imap(reductions, \(reduction, reduction_name){
    header <-
      knitr::knit_expand(
        text = "\n## {{REDUCTION_NAME}} {.tabset .tabset-fade .tabset-pills}\n",
        REDUCTION_NAME = reduction_name
      )
    each_group <-
      purrr::imap(groupings, \(grouping_var, grouping_var_name){
        subheader <-
          knitr::knit_expand(
            text = "\n### {{REDUCTION_NAME}} by {{GROUPING_VAR_NAME}} \n",
            REDUCTION_NAME = reduction,
            GROUPING_VAR_NAME = grouping_var_name
          )
        dimplot <-
          knitr::knit_expand(
            file           = here::here("analysis/templates/dimplots.Rmd"),
            REDUCTION = reduction,
            GROUPING_VAR = grouping_var
          )
        c(subheader, dimplot)
      })
    c(header, unlist(each_group))
  })
template_res <- knitr::knit_child(text = unlist(dimplots), quiet = TRUE)
cat(template_res, sep = '\n')
```

\newpage

```{r module_scores, echo = FALSE}
targets::tar_load(all_module_scores)
targets::tar_load(ifn_modules)
targets::tar_load(inflame_modules)
targets::tar_load(ldg_modules)
targets::tar_load(module_scores_mat)
targets::tar_load(study_md)
targets::tar_load(group_pal)

row_groupings  <-
  list(
    project_params[["comparison_grouping_variable"]],
    "cluster",
    "hierarchical clustering"
    )

row_groupings <- rlang::set_names(x = row_groupings, nm = row_groupings)

module_groups <-
  list(
    "Banchereau interferon modules"   = ifn_modules,
    "Banchereau inflammation modules" = inflame_modules,
    "Low-density granulocyte modules" = c("ldg_a", "ldg_b", "ldg_1.1", "ldg2.1")
  )
```

# Module score heatmaps {.tabset .tabset-fade .tabset-pills}

```{r module_score_plots, results = "asis", echo = FALSE}
modplots <-
  purrr::imap(module_groups, \(module, title){
    header <-
      knitr::knit_expand(
        text     = "\n## {{MODULE_GROUP}} scores {.tabset .tabset-fade .tabset-pills}\n",
        MODULE_GROUP = title
      )
    each_comparison <-
      purrr::imap(row_groupings, \(group, y){
        subheader <-
          knitr::knit_expand(
            text    = "\n### Sorted by {{SORTING}}\n",
            SORTING = group
          )
        ghm <-
          knitr::knit_expand(
            file            = here::here("analysis/templates/groupedComplexHeatmaps.Rmd"),
            DATASET         = "module_scores_mat",
            VARIABLE_LIST   = list(module),
            METADATA        = "study_md",
            PALETTES        = "group_pal",
            ROW_GROUPING    = group,
            ROW_ANNOTATION  = heatmap_row_annotations,
            COL_GROUPING    = "none",
            COL_ANNOTATION  = "'none'",
            SCALE           = "TRUE",
            MODULE_GROUP    = stringr::str_replace(string = title, pattern = " ", replacement = "_"),
            SAMPLE_GROUPING = stringr::str_replace(string = group, pattern = " ", replacement = "_"),
            FIG_WIDTH       = min(((1*length(module)) + 3.84), 12),
            FIG_HEIGHT      = 9
          )
        c(subheader, unlist(ghm))
      })
    c(header, unlist(each_comparison))
  })
template_res <- knitr::knit_child(text = unlist(modplots), quiet = TRUE)

cat(template_res, sep = '\n', fill = TRUE)
```

```{r}
targets::tar_load(annotated_modules)
targets::tar_load(annotated_mod_list)
targets::tar_load(banchereau_modules)
targets::tar_load(ldg_modules)
targets::tar_load(metasignature_module)
targets::tar_load(module_annotation)

annotation_list <-
  tibble::enframe(
    x     = annotated_mod_list,
    name  = "obsv",
    value = "annotation"
    ) |>
  dplyr::mutate(
    annotation =
      stringr::str_extract(
        string = annotation,
        pattern = "(?<=([[:space:]]-[[:space:]])).+"
        )
    )

module_groups <-
  list(
    "annotated Banchereau modules"    = annotation_list[['obsv']],
    "all Banchereau modules"          = names(banchereau_modules)
  )

present_module_types <-
  module_annotation |>
  dplyr::filter(module %in% names(banchereau_modules)) |>
  dplyr::pull(type) |>
  forcats::fct_drop() |>
  levels()
```

```{r, echo=FALSE, results='asis'}
groups <-
  purrr::imap(module_groups, \(module, title){
    header <-
      knitr::knit_expand(
        text     = "\n\n## Scores for {{MODULE_GROUP}} {.tabset .tabset-fade .tabset-pills}\n",
        MODULE_GROUP = title
      )
    each_comparison <-
      purrr::map(row_groupings, \(group){
        subheader <-
          knitr::knit_expand(
            text    = "\n\n### Sorted by {{SORTING}}\n\n",
            SORTING = group
          )
        heatmap <-
          knitr::knit_expand(
            file           = here::here("analysis/templates/groupedComplexHeatmaps.Rmd"),
            DATASET        = "module_scores_mat",
            VARIABLE_LIST  = list(module),
            METADATA       = "study_md",
            PALETTES       = "group_pal",
            ROW_GROUPING   = group,
            ROW_ANNOTATION = heatmap_row_annotations,
            COL_GROUPING   = "type",
            COL_ANNOTATION = "dplyr::rename(module_annotation, obsv = 'module')",
            SCALE          = "TRUE",
            FIG_WIDTH       = min(min(((1*length(module)) + 3.84), 12), 12),
            FIG_HEIGHT      = 9
          )
        c(subheader, heatmap)
      })
    c(header, unlist(each_comparison))
  })
template_res <- knitr::knit_child(text = unlist(groups), quiet = TRUE)
cat(template_res, sep = '\n')
```

## Expression of the annotated Banchereau modules {.tabset .tabset-fade .tabset-pills}

```{r}
targets::tar_load(all_module_scores_with_md)
targets::tar_load(module_comparisons_stats)
targets::tar_load(annotated_module_scores)
targets::tar_load(annotated_module_scores_pivot)
targets::tar_load(annotated_module_comparisons_stats)
targets::tar_load(comparison_groups)

module_bins <-
  split(
    x = unique(annotated_modules$module),
    f = ceiling(seq_along(unique(annotated_modules$module)) / 9)
    )
```

```{r annotated_modules_score_violins, results='asis'}
groups <-
  purrr::map(comparison_groups, \(grouping){
    header <-
      knitr::knit_expand(
        text     = "\n\n### By {{GROUPING}} {.tabset .tabset-fade .tabset-pills}\n\n",
        GROUPING = grouping
      )
    each_comparison <-
      purrr::map(module_bins, \(bin){
        knitr::knit_expand(
          file       = here::here("analysis/templates/violinPanel.Rmd"),
          EXPR_TBL   = "annotated_module_scores_pivot",
          GROUPING   = grouping,
          STATS_TBL  = "annotated_module_comparisons_stats",
          VALUES_VAR = "score",
          FACET_VAR  = "module",
          BIN        = list(bin),
          FILTER_VAR = "module"
        )
      })
    c(header, unlist(each_comparison))
  })
template_res <- knitr::knit_child(text = unlist(groups), quiet = TRUE)
cat(template_res, sep = '\n')
```

\newpage

# WGCNA

```{r include=FALSE}
targets::tar_load(wgcna_modules)
```

## Expression similarity dendrogram

```{r WGCNA_Expression_similarity_dendrogram, echo=FALSE, fig.height=6, fig.width=10}
WGCNA::plotDendroAndColors(
  dendro = wgcna_modules$dendrograms[[1]],
  colors = wgcna_modules$colors[wgcna_modules$blockGenes[[1]]],
  groupLabels = "Module colors",
  main = "Gene dendrogram and module colors in block 1",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05
)
```

## Expression of identified modules {.tabset .tabset-fade .tabset-pills}

```{r include=FALSE}
targets::tar_load(wgcna_scores_pivot)
targets::tar_load(wgcna_comparisons_stats)

module_bins <-
  split(
    x = unique(wgcna_scores_pivot$module),
    f = ceiling(seq_along(unique(wgcna_scores_pivot$module)) / 9)
    )
```

\newpage

```{r wgcna_score_violins, fig.height=8, fig.width=12, warning=FALSE, results="asis"}
groups <-
  purrr::map(comparison_groups, \(grouping){
    header <-
      knitr::knit_expand(
        text     = "\n\n### By {{GROUPING}} {.tabset .tabset-fade .tabset-pills}\n\n",
        GROUPING = grouping
      )
    each_comparison <-
      purrr::map(module_bins, \(bin){
        if (length(bin) > 1){
          knitr::knit_expand(
            file       = here::here("analysis/templates/violinPanel.Rmd"),
            EXPR_TBL   = "wgcna_scores_pivot",
            GROUPING   = grouping,
            STATS_TBL  = "wgcna_comparisons_stats",
            VALUES_VAR = "score",
            FACET_VAR  = "module",
            BIN        = list(bin),
            FILTER_VAR = "module"
          )
        } else {
          knitr::knit_expand(
              file       = here::here("analysis/templates/singleViolin.Rmd"),
              EXPR_TBL   = "wgcna_scores_pivot",
              GROUPING   = grouping,
              STATS_TBL  = "wgcna_comparisons_stats",
              VALUES_VAR = "score",
              FACET_VAR  = "module",
              BIN        = list(c(bin, "")),
              FILTER_VAR = "module"
          )
        }
      })
    c(header, unlist(each_comparison))
  })
template_res <- knitr::knit_child(text = unlist(groups), quiet = TRUE)
cat(template_res, sep = '\n')
```

\newpage

## WGCNA eigenvalues {.tabset .tabset-fade .tabset-pills}

```{r wgcna_eigenvalues, fig.height=9, fig.width=11}
targets::tar_load(wgcna_scores)

wgcna_module_colors <- paste0("ME", targets::tar_read(wgcna_module_colors))
wgcna_scores_mat <-
  module_scores_mat[stringr::str_detect(string = rownames(module_scores_mat), pattern = "^ME"),]
```

\newpage

```{r WGCNA_eigenvalues_heatmap, results='asis'}
groups <-
  purrr::map(row_groupings, \(group){
    header <-
      knitr::knit_expand(
        text    = "\n\n### Sorted by {{SORTING}}\n\n",
        SORTING = group
      )
    heatmap <-
      knitr::knit_expand(
        file           = here::here("analysis/templates/groupedComplexHeatmaps.Rmd"),
        DATASET        = "wgcna_scores_mat",
        VARIABLE_LIST  = list(wgcna_module_colors),
        METADATA       = "study_md",
        PALETTES       = "group_pal",
        ROW_GROUPING   = group,
        ROW_ANNOTATION = heatmap_row_annotations,
        COL_GROUPING   = "none",
        COL_ANNOTATION = "'none'",
        SCALE          = "FALSE",
        FIG_WIDTH      = min(((1*length(wgcna_module_colors)) + 3.84), 12),
        FIG_HEIGHT     = 6
      )
    c(header, heatmap)
  })
template_res <- knitr::knit_child(text = unlist(groups), quiet = TRUE)
cat(template_res, sep = '\n')
```


\newpage

## Gene ontology GSEA of WGCNA modules

```{r wgcna_gsea, fig.height=11, fig.width=12}
targets::tar_load(MEplotting)

module_bins <-
  split(
    x = unique(MEplotting$module),
    f = ceiling(seq_along(unique(MEplotting$module)) / 6)
  )
```

```{r wgcna_gsea_plots, results='asis'}
groups <-
  purrr::map(.x = module_bins, .f = \(bin){
    if (length(bin) > 1){
      gsea_plots <-
        knitr::knit_expand(
          file    = here::here("analysis/templates/groupedModuleGSEAPlots.Rmd"),
          DATASET = "MEplotting",
          BIN     = list(bin)
        )
    } else {
      gsea_plots <-
        knitr::knit_expand(
          file    = here::here("analysis/templates/singleModuleGSEAPlot.Rmd"),
          DATASET = "MEplotting",
          BIN     = list(c(bin, ""))
          )
    }
  })
template_res <- knitr::knit_child(text = unlist(groups), quiet = TRUE)
cat(template_res, sep = '\n')
```

\newpage

# Use of gene expression modules in classifying samples.

```{r wgcna_cluster_confusion}
targets::tar_load(module_rf_models)
targets::tar_load(wgcna_rf_models)

module_groups <-
  list(
    WGCNA      = wgcna_rf_models,
    Banchereau = module_rf_models
  )
```


```{r wgcna_cluster_importance, results='asis'}
groups <-
  purrr::imap(module_groups, \(module_group, title){
    header <-
      knitr::knit_expand(
        text     = "\n\n## {{MODULE_GROUP}} scores {.tabset .tabset-fade .tabset-pills}\n",
        MODULE_GROUP = title
      )
    each_comparison <-
      purrr::map(comparison_groups, \(group){
        subheader <-
          knitr::knit_expand(
            text    = "\n### Sorted by {{SORTING}}\n",
            SORTING = group
          )
        heatmap <-
          knitr::knit_expand(
            file      = here::here("analysis/templates/randomForestPlots.Rmd"),
            MODULE_GROUP = title,
            GROUP = group,
            CLASS_VAR = group
          )
        c(subheader, heatmap)
      })
    c(header, unlist(each_comparison))
  })
template_res <- knitr::knit_child(text = unlist(groups), quiet = TRUE)
cat(template_res, sep = '\n')

```

```{r}
sessioninfo::session_info()
```
